Left-linear rules in adhesive categories

### 2024-02-12 icalp/fusione.tex
abstract

\begin{abstract}
  When can two sequences of steps executed by a computational device
  be considered equivalent?  This is a relevant question if one
  considers concurrent and distributed systems, since the independence
  of two steps means that they could be executed in any order and
  potentially in parallel.
  % 
  We investigate the issue in the context of adhesive categories, a
  very general instance of the DPO approach to rewriting. More
  precisely, we focus on left-linear rules, which not only consume,
  preserve and generate entities, as linear rules do, but also have
  the capability of ``merging'' parts of the state.
  % 
  An apparently minimal, yet technically hard enhancement, 
  % 
  % We refine the established notion of sequential independence
  % between steps, and ...
  since the canonical notion of independence used for linear rewriting
  systems does not work out of the box.
  %
  Moreover, a standard characterisation of independence that - in the
  linear case - allows one to derive a number of properties,
  essential in the development of a theory of concurrency, no longer
  holds, and this requires a substantial revision of the
  theory.
  %
  The paper performs an in-depth study of the notion of independence,
  for left-linear rules, introduces a novel characterisation of
  independence, identifies well-behaved classes of left-linear
  rewriting systems and provides some fundamental results including a
  Church-Rosser property and the existence of canonical equivalence
  proofs for concurrent ccomputations. These results properly extends
  the class of formalisms that can be modelled in the adhesive
  framework.
\end{abstract}

>>>>

\begin{abstract}
  When can two sequences of steps executed by a computational device
  be considered equivalent?  This is a relevant question if one
  considers concurrent and distributed systems, since the independence
  of two steps means that they could be executed in any order and
  potentially in parallel.
  % 
  We investigate the issue in the context of adhesive categories, a
  very general instance of the DPO approach to rewriting. More
  precisely, we focus on left-linear rules, which not only consume,
  preserve and generate entities, as linear rules do, but also have
  the capability of ``merging'' parts of the state.
  % 
  An apparently minimal, yet technically hard enhancement, 
  % 
  % We refine the established notion of sequential independence
  % between steps, and ...
  since the canonical notion of independence used for linear rewriting
  systems does not work out of the box.
  %
  Moreover, a standard characterisation of independence that - in the
  linear case - allows one to derive a number of properties,
  essential in the development of a theory of concurrency, no longer
  holds, and this requires a substantial revision of the
  theory.
  %
  The paper performs an in-depth study of the notion of independence,
  for left-linear rules, introduces a novel characterisation of
  independence, identifies well-behaved classes of left-linear
  rewriting systems and provides some fundamental results including a
  Church-Rosser property and the existence of canonical equivalence
  proofs for concurrent computations. These results properly extends
  the class of formalisms that can be modelled in the adhesive
  framework.
\end{abstract}
